Untitled

### 2021-02-25 ms/manuscript.Rmd
Update MS, SI
- add ESM 2 (gameto dt batches)
- add time bars for gameto dt treatment, 2012 data

We selected the island of Moorea, French Polynesia (17°32' S, 149°50' W) for our survey of filmy fern ecophysiology.
Moorea is a small (135 km^2^; maximum elevation 1206 m) tropical oceanic island, located \circa{ }15 km NW of its larger sister island, Tahiti.
Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Murdock2003; @Ranker2005; @Nitta2011b; @Nitta2017; @Florence] (\figs `r figure_num("habit")`, `r s_figure_num("tree")`).
The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages [@Pryer2001a], are represented on Moorea by `r n_hymeno_sp` and `r n_trich_sp` species, respectively.
We follow the taxonomic system of @Ebihara2006 for Hymenophyllaceae.

@@

Samples were collected in the field and stored in plastic bags with a small amount of water to keep them fresh during transport to the lab.
For sporophytes, 8 individuals (one individual = single whole frond including a \circa{ }2 cm section of rhizome) were used per treatment, and all individuals for each species came from a single population.
We did not attempt to differentiate genotypes between fronds in species with long-creeping rhizomes, which often form clonal mats; hence, in some cases, individuals may be genetically identical ramets.
All experiments were initiated within 48 h of collection.
Pre-treatment maximum photochemical yield of photosystem II (F~v~/F~m~) [@Kitajima1975] was measured in fresh plants after a 10 min period of dark-adaptation using a portable mini-PAM fluorometer (Walz Gmbh, Effeltrich, Germany).
Any samples with pre-treatment F~v~/F~m~ < 0.4 were excluded from analysis, as typical values of F~v~/F~m~ in healthy plants tend to be \circa{ }0.8 [@Bjorkman1987].
Samples were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100 % RH), and water withheld for either a short (2 d) or long (15 d) interval [@Testo2013].
Conditions inside the desiccation chambers were monitored during the experiment using Track-It RH/Temp dataloggers (Monarch Instrument, Amherst, NH) logging every 30 min (2012 field season) or Hobo ProV2 dataloggers logging every 5 min (2013, 2014 field seasons).
The desiccation chambers were kept in an air-conditioned room. 
Temperature inside the chambers was mean `mean(filmy_dt_chamber$temp) %>% round(1)` ± SD `r sd(filmy_dt_chamber$temp) %>% round(2)` °C (\figs `r s_figure_num("chamber-s")`, `r s_figure_num("chamber-g")`).
Salts used for desiccation and their corresponding mean water potentials and approximate relative humidity and VPD are as follows: LiCl (-282 MPa, 18 % RH, 2.45 kPa), Mg(NO~3~)~2~ (-86 MPa, 58 % RH, 1.25 kPa), and NaCl (-38 MPa, 80 % RH, 0.60 kPa).
Following the desiccation treatment, plants were rewetted by wrapping them in wet tissues using water from a local stream.
Plants were allowed to rehydrate, and F~v~/F~m~ was again measured at 0.5 h, 24 h, and 48 h following rewetting.
Ability to recover from desiccation was quantified by comparing pre- and post-treatment F~v~/F~m~ [@WatkinsJr.2007c].

@@

Because field-collected gametophytes could not be identified to species prior to DNA extraction, no planned replication by species was possible.
We therefore applied the same treatment for all gametophytes (2 d at -86 MPa), and verified species of each individual after the experiment using the DNA barcode approach described above.
Thus, experiments involving gametophytes were sampled many times for some species whereas other species were not.
Relative water content was not calculated for gametophytes, as these were too small to measure accurately with the balances available.

@@

Relative humidity in the growth chambers was high when samples were first inserted, then gradually decreased to the target value maintained by each drying salt (\figs{ }`r s_figure_num("chamber-s")`, `r s_figure_num("chamber-g")`).
Brief fluctuations to higher humidities were observed when chambers were opened to insert or remove samples (`r s_figure("chamber-s")`).

>>>>

We selected the island of Moorea, French Polynesia (17°32' S, 149°50' W) for our survey of filmy fern ecophysiology.
Moorea is a small (135 km^2^; maximum elevation 1206 m) tropical oceanic island, located \circa{ }15 km NW of its larger sister island, Tahiti.
Despite its small size, it hosts a range of habitats including coastal strand, low elevation rainforest, and high elevation cloud forest.
A total of `r n_filmy_sp` filmy fern species are known from Moorea, including multiple terrestrial, saxicolous, and epiphytic species [@Murdock2003; @Ranker2005; @Nitta2011b; @Nitta2017; @Florence] (\figs `r figure_num("habit")`, `r s_figure_num("tree")`; all supplemental figures available in Online Resource 1).
The two major clades of filmy ferns, the hymenophylloid and trichomanoid lineages [@Pryer2001a], are represented on Moorea by `r n_hymeno_sp` and `r n_trich_sp` species, respectively.
We follow the taxonomic system of @Ebihara2006 for Hymenophyllaceae.

@@

Samples were collected in the field and stored in plastic bags with a small amount of water to keep them fresh during transport to the lab.
For sporophytes, 8 individuals (one individual = single whole frond including a \circa{ }2 cm section of rhizome) were used per treatment, and all individuals for each species came from a single population.
We did not attempt to differentiate genotypes between fronds in species with long-creeping rhizomes, which often form clonal mats; hence, in some cases, individuals may be genetically identical ramets.
All experiments were initiated within 48 h of collection.
Pre-treatment maximum photochemical yield of photosystem II (F~v~/F~m~) [@Kitajima1975] was measured in fresh plants after a 10 min period of dark-adaptation using a portable mini-PAM fluorometer (Walz Gmbh, Effeltrich, Germany).
Any samples with pre-treatment F~v~/F~m~ < 0.4 were excluded from analysis, as typical values of F~v~/F~m~ in healthy plants tend to be \circa{ }0.8 [@Bjorkman1987].
Samples were then transferred to desiccation chambers containing saturated salts at three different desiccation intensities or a control treatment with moist tissues (100 % RH), and water withheld for either a short (2 d) or long (15 d) interval [@Testo2013].
Conditions inside the desiccation chambers were monitored during the experiment using Track-It RH/Temp dataloggers (Monarch Instrument, Amherst, NH) logging every 10 min (2012 field season) or Hobo ProV2 dataloggers logging every 5 min (2013, 2014 field seasons).
The desiccation chambers were kept in an air-conditioned room. 
Temperature inside the chambers was mean `mean(filmy_dt_chamber$temp) %>% round(1)` ± SD `r sd(filmy_dt_chamber$temp) %>% round(2)` °C (\figs `r s_figure_num("chamber")`).
Salts used for desiccation and their corresponding mean water potentials and approximate relative humidity and VPD are as follows: LiCl (-282 MPa, 18 % RH, 2.45 kPa), Mg(NO~3~)~2~ (-86 MPa, 58 % RH, 1.25 kPa), and NaCl (-38 MPa, 80 % RH, 0.60 kPa).
Following the desiccation treatment, plants were rewetted by wrapping them in wet tissues using water from a local stream.
Plants were allowed to rehydrate, and F~v~/F~m~ was again measured at 0.5 h, 24 h, and 48 h following rewetting.
Ability to recover from desiccation was quantified by comparing pre- and post-treatment F~v~/F~m~ [@WatkinsJr.2007c].

@@

Because field-collected gametophytes could not be identified to species prior to DNA extraction, no planned replication by species was possible.
We therefore applied the same treatment for all gametophytes (2 d at -86 MPa) in batches of individuals (`r s_table("gameto-batches")`), and verified the species of each individual after the experiment using the DNA barcode approach described above.
Thus, experiments involving gametophytes were sampled many times for some species whereas other species were not.
Relative water content was not calculated for gametophytes, as these were too small to measure accurately with the balances available.

@@

Relative humidity in the growth chambers was high when samples were first inserted, then gradually decreased to the target value maintained by each drying salt (\figs{ }`r s_figure_num("chamber")`).
Brief fluctuations to higher humidities were observed when chambers were opened to insert or remove samples (`r s_figure("chamber-s")`).
